Untitled

### 2023-04-10 Semester II/Research Method/Manuskrip ke Software X/manuskrip versi simpel pdf/template.Rmd
update

---
title: An R-base program to build e-commerce product discount portfolio using spiral dynamic optimization algorithm
author:
  - name: Mohammad Rizka Fadhli
    email: 20921004@mahasiswa.itb.ac.id
    correspondingauthor: true
    affiliation: lala
  - name: Saladin Uttunggadewa
    affiliation: lili
  - name: Rieske Hadianti
    affiliation: lulu
  - name: Sri Redjeki Pudjaprasetya
    affiliation: lili
address:
  - code: lala
    address: Magister of Computational Sciences Program, Faculty of Mathematics and Natural Sciences, Institut Teknologi Bandung
  - code: lili
    address: Faculty of Mathematics and Natural Sciences, Institut Teknologi Bandung
  - code: lulu
    address: Center for Mathematical Modeling and Simulation, Institut Teknologi Bandung
abstract: |
  The development of the digital economy market in Indonesia is growing every year. As a result, marketplaces have to compete with each other to get consumers. One of the marketplace startup in Indonesia has strategy to provide additional discounts on their listed products. Product discount portfolio problem is a Binary Linear Programming (BLP) problem with a decision variable in the form of a binary number that states whether a product is eligible for a discount or not. With many decision variables involved, we developed metaheuristic approaches Spiral Dynamic Optimization Algorithm (SDOA) to solve the optimization problem. The SDOA solution is proven to be more optimal (generates higher revenur) than the existing startup solution by average $9.8 \%$.
keywords: 
  - binary linear programming
  - spiral dynamic optimization algorithm
  - R programming
  - e-commerce
journal: "Software X"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: mybibfile.bib
linenumbers: false
numbersections: true
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

@@

One of the crucial processes in SDOA is to rotate candidate solutions using a rotation matrix. In this case, one candidate solution is a vector of size 1 $\times$ 100 while the rotation matrix is of size 100 $\times$ 100. This process is quite computationally burdensome, especially since the number of candidate solutions used is relatively large. Therefore, we need to modify this rotation process by using matrix multiplication partitioning in linear algebra.

@@

The algorithm for solving optimization problems using modified SDOA was made in R programming language. As a form of evaluation of the algorithm, we compared the SDOA solution with:

>>>>

---
title: "Untitled"
output: word_document
date: "2023-03-13"
bibliography: references.bib
csl: csl new.csl
---

@@

One of the crucial processes in SDOA is to rotate candidate solutions using a rotation matrix. For instance, this is R program to create roration matrix using input $n$ dimension and $\theta$ as rotation angle.

```
rot_mat_creator = function(theta,n){
  # initiate zero matrix n x n
  temp_mat          = matrix(0,ncol = n,nrow = n)
  # put `1` into matrix diagonal
  diag(temp_mat)    = 1
  # initiate matrix rotation
  mat_rot           = temp_mat
  # iterate to calculate matrikx element
  for(i in 1:(n-1)){
    for(j in 1:i){
      temp          = temp_mat
      idx           = n-i
      idy           = n+1-j
      temp[idx,idx] = cos(theta)
      temp[idx,idy] = -sin(theta)
      temp[idy,idx] = sin(theta)
      temp[idy,idy] = cos(theta)
      mat_rot       = mat_rot %*% temp
    }
  }
  return(mat_rot)
}
```

In this case, one candidate solution is a vector of size 1 $\times$ 100 while the rotation matrix is of size 100 $\times$ 100. This process is quite computationally burdensome, especially since the number of candidate solutions used is relatively large. Therefore, we need to modify this rotation process by using matrix multiplication partitioning in linear algebra.

@@

The algorithm for solving optimization problems using modified SDOA was made in R programming language only using base library and dplyr [@dplyr] library for data preparation and data wrangling. As a form of evaluation of the algorithm, we compared the SDOA solution with:

### 2023-04-10 Semester II/Research Method/Manuskrip ke Software X/manuskrip versi simpel pdf/template.tex
update

\documentclass[preprint, 3p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{Software X} % Sets Journal name

\usepackage{lineno} % add

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={An R-base program to build e-commerce product discount portfolio using spiral dynamic optimization algorithm},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\begin{document}

\begin{frontmatter}

  \title{An R-base program to build e-commerce product discount
portfolio using spiral dynamic optimization algorithm}
    \author[lala]{Mohammad Rizka Fadhli%
  \corref{cor1}%
  }
   \ead{20921004@mahasiswa.itb.ac.id} 
    \author[lili]{Saladin Uttunggadewa%
  %
  }

    \author[lulu]{Rieske Hadianti%
  %
  }

    \author[lili]{Sri Redjeki Pudjaprasetya%
  %
  }

      \affiliation[lala]{Magister of Computational Sciences Program,
Faculty of Mathematics and Natural Sciences, Institut Teknologi Bandung}
    \affiliation[lili]{Faculty of Mathematics and Natural Sciences,
Institut Teknologi Bandung}
    \affiliation[lulu]{Center for Mathematical Modeling and Simulation,
Institut Teknologi Bandung}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  The development of the digital economy market in Indonesia is growing
  every year. As a result, marketplaces have to compete with each other
  to get consumers. One of the marketplace startup in Indonesia has
  strategy to provide additional discounts on their listed products.
  Product discount portfolio problem is a Binary Linear Programming
  (BLP) problem with a decision variable in the form of a binary number
  that states whether a product is eligible for a discount or not. With
  many decision variables involved, we developed metaheuristic
  approaches Spiral Dynamic Optimization Algorithm (SDOA) to solve the
  optimization problem. The SDOA solution is proven to be more optimal
  (generates higher revenur) than the existing startup solution by
  average \(9.8 \%\).
  \end{abstract}
    \begin{keyword}
    binary linear programming \sep spiral dynamic optimization
algorithm \sep R programming \sep 
    e-commerce
  \end{keyword}

 \end{frontmatter}

@@

One of the top marketplace in Indonesia use a product discount strategy
to win the competition. Each listed product on the marketplace has an
initial price defined by the seller. Marketplace can intervene in its
price indirectly by providing discounts, so consumers pay lower than the
initial price. A study at an online retailer in China shows an influence
between product discounts and consumer purchase behavior such as
purchase incident, purchase quantity, and spending. Especially in
specific discount percentage ranges \citep{cny}. The pricing and
discount aim to attract consumers to buy products in the marketplace for
a certain period. One of the mathematical models that can be used to
determine prices is the Price Elasticity Model, a causality model
between price and demand. A study comparing the price elasticity model
with linear, polynomial, and exponential bases found that the linear
model is the most commonly used, and it is easier to interpret
\citep{linear_reg}.

The problem is determining which products need to be given an additional
discount. This problem is one of the binary linear programming (BLP)
forms: choosing the right product to maximize sales with a constraint
limited discount budget in a certain period. Binary programming is a
form of an optimization method in which the variables involved are
binary numbers. The main characteristic of BLP is the matching process
between indexes \citep{lieberman}. One of the studies related to BLP is
the matching process to determine the placement of phasor measuring
units in a power system \citep{suresh}. Another study used BLP to
optimize the selection of 100 marketing channels and activities against
millions of customers \citep{ieee}. The goal is to optimize the
marketing messages delivery so that sales increase. Business-related and
consumer-centric optimization problems have not been widely published in
journals.

At least we can use two methods to solve BLP: the exact method and the
meta-heuristic method. Meta-heuristic methods are developing rapidly
that several algorithms can quickly solve optimization problems. The
meta-heuristic algorithm such as SDOA can solve BLP in a small number of
decision variables involved. SDOA is an algorithm inspired by spiral
motion in natural events. The rotation matrix is an algebraic principle
closely related to SDOA \citep{tamura}. SDOA is proven to be used to
solve BLP problems by modifying the existing objective function and
constraints \citep{kun}. The basic idea is to turn a constrained problem
into an unconstrained problem by creating a penalty function.

In another study, researchers made an adaptive linear SDOA by changing
the value of the contraction constant into a specific function that
depends on the function value of each candidate solution \citep{sdoa2}.
The latest study in 2022 conducted a thorough review, including making
several improvements to SDOA \citep{sdoa3}, such as:

@@

One of the crucial processes in SDOA is to rotate candidate solutions
using a rotation matrix. In this case, one candidate solution is a
vector of size 1 \(\times\) 100 while the rotation matrix is of size 100
\(\times\) 100. This process is quite computationally burdensome,
especially since the number of candidate solutions used is relatively
large. Therefore, we need to modify this rotation process by using
matrix multiplication partitioning in linear algebra.

@@

The algorithm for solving optimization problems using modified SDOA was
made in R programming language. As a form of evaluation of the
algorithm, we compared the SDOA solution with:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The existing solution from the marketplace.
\item
  The exact solution uses the simplex method created using the R
  programming language with \texttt{library(ompr)} \citep{ompr}.
\end{enumerate}

@@

\renewcommand\refname{REFERENCES}
\bibliography{mybibfile.bib}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Untitled},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Untitled}
\author{}
\date{\vspace{-2.5em}2023-03-13}

\begin{document}
\maketitle

@@

One of the top marketplace in Indonesia use a product discount strategy
to win the competition. Each listed product on the marketplace has an
initial price defined by the seller. Marketplace can intervene in its
price indirectly by providing discounts, so consumers pay lower than the
initial price. A study at an online retailer in China shows an influence
between product discounts and consumer purchase behavior such as
purchase incident, purchase quantity, and spending. Especially in
specific discount percentage ranges {[}1{]}. The pricing and discount
aim to attract consumers to buy products in the marketplace for a
certain period. One of the mathematical models that can be used to
determine prices is the Price Elasticity Model, a causality model
between price and demand. A study comparing the price elasticity model
with linear, polynomial, and exponential bases found that the linear
model is the most commonly used, and it is easier to interpret {[}2{]}.

The problem is determining which products need to be given an additional
discount. This problem is one of the binary linear programming (BLP)
forms: choosing the right product to maximize sales with a constraint
limited discount budget in a certain period. Binary programming is a
form of an optimization method in which the variables involved are
binary numbers. The main characteristic of BLP is the matching process
between indexes {[}3{]}. One of the studies related to BLP is the
matching process to determine the placement of phasor measuring units in
a power system {[}4{]}. Another study used BLP to optimize the selection
of 100 marketing channels and activities against millions of customers
{[}5{]}. The goal is to optimize the marketing messages delivery so that
sales increase. Business-related and consumer-centric optimization
problems have not been widely published in journals.

At least we can use two methods to solve BLP: the exact method and the
meta-heuristic method. Meta-heuristic methods are developing rapidly
that several algorithms can quickly solve optimization problems. The
meta-heuristic algorithm such as SDOA can solve BLP in a small number of
decision variables involved. SDOA is an algorithm inspired by spiral
motion in natural events. The rotation matrix is an algebraic principle
closely related to SDOA {[}6{]}. SDOA is proven to be used to solve BLP
problems by modifying the existing objective function and constraints
{[}7{]}. The basic idea is to turn a constrained problem into an
unconstrained problem by creating a penalty function.

In another study, researchers made an adaptive linear SDOA by changing
the value of the contraction constant into a specific function that
depends on the function value of each candidate solution {[}8{]}. The
latest study in 2022 conducted a thorough review, including making
several improvements to SDOA {[}9{]}, such as:

@@

One of the crucial processes in SDOA is to rotate candidate solutions
using a rotation matrix. For instance, this is R program to create
roration matrix using input \(n\) dimension and \(\theta\) as rotation
angle.

\begin{verbatim}
rot_mat_creator = function(theta,n){
  # initiate zero matrix n x n
  temp_mat          = matrix(0,ncol = n,nrow = n)
  # put `1` into matrix diagonal
  diag(temp_mat)    = 1
  # initiate matrix rotation
  mat_rot           = temp_mat
  # iterate to calculate matrikx element
  for(i in 1:(n-1)){
    for(j in 1:i){
      temp          = temp_mat
      idx           = n-i
      idy           = n+1-j
      temp[idx,idx] = cos(theta)
      temp[idx,idy] = -sin(theta)
      temp[idy,idx] = sin(theta)
      temp[idy,idy] = cos(theta)
      mat_rot       = mat_rot %*% temp
    }
  }
  return(mat_rot)
}
\end{verbatim}

In this case, one candidate solution is a vector of size 1 \(\times\)
100 while the rotation matrix is of size 100 \(\times\) 100. This
process is quite computationally burdensome, especially since the number
of candidate solutions used is relatively large. Therefore, we need to
modify this rotation process by using matrix multiplication partitioning
in linear algebra.

@@

The algorithm for solving optimization problems using modified SDOA was
made in R programming language only using base library and dplyr
{[}10{]} library for data preparation and data wrangling. As a form of
evaluation of the algorithm, we compared the SDOA solution with:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The existing solution from the marketplace.
\item
  The exact solution uses the simplex method created using the R
  programming language with \texttt{library(ompr)} {[}11{]}.
\end{enumerate}

@@

\hypertarget{references}{%
\section*{REFERENCES}\label{references}}
\addcontentsline{toc}{section}{REFERENCES}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-cny}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{Liu, H., Lobschat, L., Verhoef, P., \& Zhao, H. (2020).
The effect of permanent product discounts and order coupons on purchase
incidence, purchase quantity, and spending. \emph{Journal of Retailing}.
\url{https://doi.org/10.1016/j.jretai.2020.11.007}}

\leavevmode\vadjust pre{\hypertarget{ref-linear_reg}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{Hajdinjak, M. (2020). Functions with linear price
elasticity for forecasting demand and supply. \emph{The B.E. Journal of
Theoretical Economics}. \url{https://doi.org/10.1515/bejte-2017-0015}}

\leavevmode\vadjust pre{\hypertarget{ref-lieberman}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{Hillier, F. S., \& Lieberman, G. J. (2001).
\emph{Introduction to operations research} (7th ed.). New York, US:
McGraw Hill.}

\leavevmode\vadjust pre{\hypertarget{ref-suresh}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Billakanti, S., \& Venkaiah, Ch. (2014). An effective
binary integer linear programmed approach for optimal placement of PUMs
in power systems. \emph{IEEE: International Conference on Smart ELectric
Grid}. \url{https://doi.org/10.1109/ISEG.2014.7005593}}

\leavevmode\vadjust pre{\hypertarget{ref-ieee}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{Bigler, T., Baumann, P., \& Kammermann, M. (2019).
Optimizing customer assignments to direct marketing activities: A binary
linear programming formulation. \emph{IEEE International Conference on
Industrial Engineering and Engineering Management (IEEM)}.
\url{https://doi.org/10.1109/IEEM44572.2019.8978863}}

\leavevmode\vadjust pre{\hypertarget{ref-tamura}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{Tamura, K., \& Yasuda, K. (2011). Spiral dynamics
inspired optimization. \emph{Journal of Advanced Computational
Intelligence and Intelligent Informatics}.
\url{https://doi.org/10.20965/jaciii.2011.p1116}}

\leavevmode\vadjust pre{\hypertarget{ref-kun}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{Kania, A., \& Sidarto, K. A. (2016). Solving mixed
integer nonlinear programming problems using spiral dynamics
optimization algorithm. \emph{AIP Publishing}.
\url{https://doi.org/10.1063/1.4942987}}

\leavevmode\vadjust pre{\hypertarget{ref-sdoa2}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{Nasir, A. N. K., Ismail, R. M. T. R., \& Tokhi, M. O.
(2016). Adaptive spiral dynamics metaheuristic algorithm for global
optimisation with application to modelling of a flexible system.
\emph{Applied Mathematical Modelling}.
\url{https://doi.org/10.1016/j.apm.2016.01.002}}

\leavevmode\vadjust pre{\hypertarget{ref-sdoa3}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{Omar, M. B., Bingi, K., Prusty, B. R., \& Ibrahim, R.
(2022). Recent advances and applications of spiral dynamics optimization
algorithm: A review. \emph{MDPI, Basel, Switzerland}.
\url{https://doi.org/10.3390/fractalfract6010027}}

\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{Wickham, H., François, R., Henry, L., Müller, K., \&
Vaughan, D. (2023). \emph{Dplyr: A grammar of data manipulation}.}

\leavevmode\vadjust pre{\hypertarget{ref-ompr}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{Schumacher, D. (2022). \emph{Ompr: Model and solve mixed
integer linear programs}. Retrieved from
\url{https://github.com/dirkschumacher/ompr}}

\end{CSLReferences}
